Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-09-22 writeup/rocha_etal.tex
Adds small comment about spectral differentiation

\begin{abstract}
    We provide model-based evidence of
    vigorous modulation of near-surface dynamics at submesoscales
    (10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
    are prone to baroclinic instabilities, and submesoscale turbulence
    prevails in late winter/early spring. Summertime
    re-stratification weakens submesoscale turbulence and enhances submesoscale inertia-gravity
     waves near the surface. In the Kuroshio Extension,
    inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
    sea-surface height (SSH) variance in late summer/early fall, with implications for
    the accuracy of surface geostrophic currents diagnosed from SSH measured by
    high-resolution satellite altimeters.
\end{abstract}

@@

% now SWOT
The characterization of the partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality has immediate applications for the planning
of the Surface and Water Ocean Topography (SWOT) satellite mission.
SWOT will provide
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}. A
challenge for SWOT is the presence of ageostrophic flows that straddle the submesoscales and project
onto SSH \citep[e.g., ][]{richman_etal2012} thus making submesoscale geostrophic
velocities difficult to extract from SSH.

@@

\section{Bulk statistics of the surface lateral velocity gradient tensor}
To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
\begin{equation}
\left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right]
\end{equation}
using a centered
second-order finite difference scheme. We then diagnose
the vertical vorticity
\begin{equation}
\zeta \equiv v_x - u_y\, ,
\end{equation}
lateral rate of strain
\begin{equation}
  \alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}\,,\\
\end{equation}
and horizontal divergence
\begin{equation}
\delta \equiv u_x + v_y\, .
\end{equation}
Because vortices, fronts, and IGWs are ubiquitous at submesoscales, second-order
statistics of vorticity, strain, and divergence are apt descriptors of these
fine lateral scales \citep[e.g.,][]{shcherbina_etal2013}.

>>>>

\begin{abstract}
    We provide evidence  of
    vigorous modulation of near-surface dynamics at submesoscales
    (10-100 km) in the Kuroshio Extension. The analysis uses outputs of general
    circulation models spun-up from a state estimate and forced by high-frequency
    winds, surface fluxes, and tides.
    Consistent with recent studies, deep late-winter mixed layers
    are prone to shallow baroclinic instabilities and submesoscale turbulence
    prevails in late winter/early spring. Summertime
    re-stratification, however, weakens submesoscale turbulence and enhances submesoscale inertia-gravity
     waves near the surface. In the Kuroshio Extension,
    inertia-gravity waves dominate the submesoscale surface kinetic energy and
    sea-surface height (SSH) variance in late summer/early fall, with implications for
    the accuracy of surface geostrophic currents diagnosed from SSH measured by
    high-resolution satellite altimeters.
\end{abstract}

@@

% now SWOT
The partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality have immediate applications for the planning
of the Surface and Water Ocean Topography (SWOT) satellite mission.
SWOT will provide
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}. A
challenge --- and opportunity ---  for SWOT is the presence of ageostrophic flows
that straddle the submesoscales and project
onto SSH \citep[e.g., ][]{richman_etal2012} thus making submesoscale geostrophic
velocities difficult to extract from SSH.

@@

\section{Bulk statistics of the surface lateral velocity gradient tensor}
To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
\begin{equation}
\left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right]
\end{equation}
using a centered
second-order finite difference scheme --- we obtain qualitatively similar results
using spectral differentiation of velocity fields periodized with reflections;
statistics obtained via spectral differentiation are 10-15$\%$ larger.
We then diagnose
the vertical vorticity
\begin{equation}
\zeta \equiv v_x - u_y\, ,
\end{equation}
lateral rate of strain
\begin{equation}
  \alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}\,,\\
\end{equation}
and horizontal divergence
\begin{equation}
\delta \equiv u_x + v_y\, .
\end{equation}
Because vortices, fronts, and IGWs are ubiquitous at submesoscales, second-order
statistics of vorticity, strain, and divergence are apt descriptors of these
fine lateral scales \citep[e.g.,][]{shcherbina_etal2013}.
